Automatic Matching of Bullet Lands

### 2016-10-26 imaging-paper.Rnw
No more ballistic imaging

The feasibility of creating a database of ballistic images that could be used to identify guns used in crimes was evaluated in a 2008 report~\citep{nap:2008} by the National Research Council. The evaluation was investigating the scalability of NIBIN (National Integrated Ballistic Information Network), which uses proprietary matching algorithms provided by IBIS. The bottom line of the report was that in spite of the many technical and practical hurdles, solutions to all but one problem could be found. The problem that remained is that statistically, the quality of the matching algorithm (in this case, of breech-face marks and firing pin impressions) could not withstand a hugely increased number of records while still maintaining a reasonable workload for forensic examiners, who have to examine possible matches suggested by the system. 
The findings of the NRC report on ballistic imaging are based on two-dimensional greyscale images, which the committee argued were not reliable enough for distinguishing between fine marks. This finding coincides with the assessment by \citet{dekinder:2004} based on the IBIS Heritage system. A further re-assessment by \citet{deceuster:2015} came to the same conclusions based on the EvoFinder system. 
The NRC report also found that results from 2D images can be improved when matches are based on 3D images. This is consistent with the importance of features found here: out of the top five features (see Figure~\ref{fig:importance}), only the total number of matches and mismatches are available for a match based on 2D features.

>>>>

The feasibility of creating a database of images that could be used to identify guns used in crimes was evaluated in a 2008 report~\citep{nap:2008} by the National Research Council. The evaluation was investigating the scalability of NIBIN (National Integrated Ballistic Information Network), which uses proprietary matching algorithms provided by IBIS. The bottom line of the report was that in spite of the many technical and practical hurdles, solutions to all but one problem could be found. The problem that remained is that statistically, the quality of the matching algorithm (in this case, of breech-face marks and firing pin impressions) could not withstand a hugely increased number of records while still maintaining a reasonable workload for forensic examiners, who have to examine possible matches suggested by the system. 
The findings of the NRC report on imaging are based on two-dimensional greyscale images, which the committee argued were not reliable enough for distinguishing between fine marks. This finding coincides with the assessment by \citet{dekinder:2004} based on the IBIS Heritage system. A further re-assessment by \citet{deceuster:2015} came to the same conclusions based on the EvoFinder system. 
The NRC report also found that results from 2D images can be improved when matches are based on 3D images. This is consistent with the importance of features found here: out of the top five features (see Figure~\ref{fig:importance}), only the total number of matches and mismatches are available for a match based on 2D features.
